---
title: "ProjectCheckpoint3"
format: html
editor: visual
---

## Data Summaries Using Existing Columns

# Data Summary 1

Which parent companies contribute the most total plastic waste across all countries, and what share of the total recorded waste does each company account for?

```{r}

overall_total <- sum(plastics$grand_total, na.rm = TRUE)

company_share <- function(x, overall_total) {
  sum(x, na.rm = TRUE) / overall_total
}

company_summary <- plastics |>
  filter(!is.na(parent_company),
         parent_company != "Grand Total",
         parent_company != "null",
         parent_company != "NULL",
         parent_company != "Unbranded") |>
  group_by(parent_company) |>
  summarize(
    total_plastic_waste = sum(grand_total, na.rm = TRUE),
    share_of_total = company_share(grand_total, overall_total),
    .groups = "drop"
  ) |>
  arrange(desc(total_plastic_waste)) |>
  slice_head(n = 10)

company_summary

```

# Data Summary 2

Which countries had clean up efforts that found the largest grand total of plastics in 2019 and 2020?

```{r}

## Uses iteration and a custom function
total_plastics <- function(input_country){
  
  grand_total_plastics <- plastics |>
    filter(country == input_country) |>
    distinct(year, grand_total) |>
    pull(grand_total)
  
  sum(grand_total_plastics, na.rm = TRUE)
}

plastics |>
  distinct(country) |>
  filter(country != "EMPTY") |>
  mutate(
    plastic_total = map_int(country, total_plastics)
    ) |>
  slice_max(plastic_total, n = 10)

```

## Data Summaries That Modify Columns

# Data Summary 3

Which countries have the highest average amount of plastic collected per cleanup event?

```{r}

plastic_per_event <- function(input_country) {
  plastics |>
    filter(country == input_country) |>
    distinct(year, grand_total, num_events) |>
    summarise(
      total_plastic = sum(grand_total, na.rm = TRUE),
      total_events = sum(num_events, na.rm = TRUE)
    ) |>
    mutate(plastic_per_event = total_plastic / total_events) |>
    pull(plastic_per_event)
}

plastic_event_summary <- plastics |>
  distinct(country) |>
  mutate(plastic_per_event = map_dbl(country, plastic_per_event)) |>
  slice_max(plastic_per_event, n = 10)

plastic_event_summary

```

# Data Summary 4

Which nations had the highest average participation for cleanup events in 2019 and 2020?

```{r}

cleanup_totals <- function(input_country) {
  plastics |>
    filter(country == input_country) |>
    distinct(num_events, volunteers) |>
    mutate(total_vol = sum(volunteers, na.rm = T),
         total_events = sum(num_events, na.rm = T)) |>
    distinct(total_vol, total_events) |>
    mutate(vol_per_event = total_vol / total_events) |>
    pull(vol_per_event)
}

plastics |>
  distinct(country) |>
  mutate(
    avg_participation = map_dbl(country, cleanup_totals),
    rounded_participation = round(avg_participation)
  ) |>
  slice_max(avg_participation, n = 20)

```

```{r}

library(tidyverse)
library(glue)
library(jsonlite)
library(httr)
library(readr)

plastics <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-01-26/plastics.csv')

```

## Create a function to pull the GDP of countries from the API

```{r}

my_key <- "CqRAaU4fKJAq5QtTUE57cQ5weIqdvxyC3TMQT1rY
"

get_gdp <- function(country_code, key = my_key) {
  
  url <- glue("https://api.api-ninjas.com/v1/gdp?country={country_code}")
  
  response <- GET(
    url,
    add_headers("X-Api-Key" = key)
  )
  
  data <- fromJSON(content(response, "text", encoding = "UTF-8"))
  
  if (length(data) == 0) {
    return(NA)
  } else {
    return(
      data |>
        filter(year %in% c(2019, 2020))
    )
  }
}

```

## Create a list of the top 10 countries with the highest total plastics

```{r}

library(countrycode)

top_plastic_countries <- plastics |>
  distinct(country) |>
  filter(!country %in% c("EMPTY", "Taiwan_ Republic of China (ROC)")) |>  ## Dropping Taiwan because of its contested status in the world (some supplemental data does not include Taiwan)
  mutate(
    plastic_total = map_int(country, total_plastics),
    country_code = countrycode(country, origin = "country.name", destination = "iso3c")
  ) |>
  filter(!is.na(country_code)) |> ## Dropping Nigeria's duplicate
  filter(plastic_total != 61123) |> ## Added security for removing the specific duplicate
  slice_max(plastic_total, n = 10) |>
  pull(country_code)
Sys.sleep(20)


```

## Create a list of the countries with their GDP

```{r}

gdp_list <- top_plastic_countries |>
  map(get_gdp)

```

## Double check for NAs

```{r}

gdp_list |>
  is.na() |>
  sum()

```

## Turn this to a tibble to join with original dataframe

```{r}

gdp_data <- gdp_list |>
  keep(is.data.frame) |>
  bind_rows() |>
  rename(country_code = country)

gdp_data |>
  pull(country_code)

write.csv(gdp_data, "gdp.csv", row.names = FALSE)

```

```{r}

gdp_df <- read.csv("gdp.csv")

plastics_top <- plastics |>
  filter(country != "Nigeria") |>
  filter(!country %in% c("EMPTY", "Taiwan_ Republic of China (ROC)")) |>
  mutate(
    country_code = countrycode(country, origin = "country.name", destination = "iso3c")
  ) |>
  filter(country_code %in% top_plastic_countries) |>
  left_join(gdp_df)

head(plastics_top)

```

## Data found from https://github.com/datasets/world-development-indicators/tree/main/indicators/eg.fec.rnew.zs

```{r}

setwd("~/r")
renewable_data <- read.csv("renewable-energy.csv")

plastics_top <- plastics_top |>
  left_join(renewable_data |>
              filter(Year %in% c(2019, 2020),
                     Country.Code %in% plastics_top$country_code), 
            by = c("country_code" = "Country.Code", 
                   "year" = "Year")) |>
  select(-Country.Name)

```

## What is the relationship between total plastic was and percentage of renewable energy, for the top 10 countries in total plastic waste ?

```{r}

renewable_summary <- plastics_top |>
  filter(country_code %in% top_plastic_countries) |>
  distinct(country, country_code, year, grand_total, Value) |>
  
  group_by(country, country_code) |>
  summarize(
    total_waste = sum(grand_total, na.rm = TRUE),
    renewable_pct = mean(Value, na.rm = TRUE),
    .groups = "drop"
  )

renewable_summary <- final_data |>
  filter(country != "Ecuador")

renewable_summary

```

# Plastic Waste vs. Renewable Energy Percentage by Country

```{r}

ggplot(final_data, aes(x = total_waste, y = renewable_pct)) +
  geom_point(size = 3, color = "black") +
  
  geom_text_repel(
    aes(label = country_code),
    size = 4,
    box.padding = 0.5,
    point.padding = 0.5
  ) +
  
  labs(
    title = "Plastic Waste vs Renewable Energy by Country",
    subtitle = "Top 10 Countries by Total Plastic Waste",
    x = "Total Plastic Waste",
    y = "Renewable Energy (% of total energy consumption)"
  ) +
  
  theme_classic(base_size = 13) +
  
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12),
    plot.margin = margin(15, 15, 15, 15),
    axis.title = element_text(size = 12)
  )

```
